Diagnostic Plots

### 2023-11-25 pre_print_manuscript.qmd
Linkked to diagnostic plots in manuscript

---
articletitle: |
  Self-talk interventions and sport/motor performance: An updated systematic review and Bayesian meta-analysis
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
author:
  - name: Hannah Corcoran
    affiliations:
      - ref: 1
  - name: James Steele
    affiliations:
      - ref: 1
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james.steele@solent.ac.uk
affiliations:
      - id: 1
        name: Department of Sport and Health, Solent University, UK
abstract: |
  Self-talk has been researched as an aspect of mental preparation for performance in sports and other motor tasks since the late 1980s. In 2011 Hatzigeorgiadis and colleagues systematically reviewed and meta-analysed three decades of self-talk intervention research including 32 studies. Yet, despite the general proliferation of meta-analyses, this topic has not been meta-analysed in the decade since their review which, at least for the main effect, provided a reasonably precise standardised mean difference estimate (0.48 [95% confidence interval: 0.38, 0.58]). Several further studies on self-talk interventions have been conducted in that time and it is of interest to explore what additional evidence they offer regarding the effects of self-talk interventions on sport/motor performance. Bayesian approaches are well positioned to explore how additional evidence changes our understanding of an effect; to see whether it has changed in sign, magnitude, or precision, or whether further research has largely been a ‘waste’. Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. Informative priors were taken directly from Hatzigeorgiadis et al. A total of 34 studies providing 128 effects nested in 64 groups across experiments 42 were included in the final updated meta-analysis representing data from 18761 participants. The overall posterior pooled estimate for the standardised mean difference was almost exactly the same as the prior: 0.47 [95% quantile interval: 0.39, 0.56]. Bayes factors were calculated for a range of effect sizes and indicated that the included studies largely reflected ‘Weak’ evidence against effects ranging from 0.30 to 0.59, and only provided ‘Decisive’ evidence or greater against more extreme effects: either very small (i.e., <0.02) or large (i.e., >0.81). Results were largely similar for all moderator analyses which were updated too; either providing relatively weak evidence against effects found in the previous meta-analysis or evidence suggesting smaller effects for certain moderators. The findings of our updated Bayesian meta-analyses reiterate the positive effect of self-talk interventions on sport/motor performance. However, they also suggest that cumulatively the past decade and more of research has done little to further our understanding of these effects. Considering the limited resources and time for conducting research, it may be worth moving onto to other more pertinent questions regarding psychological constructs impact upon sport/motor performance.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [psychological intervention, cumulative evidence, research waste] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/dqwh5/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: references.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

@@

As we were interested in determining how much the new evidence produced since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] had updated our understanding of the effects of self-talk interventions, we fit each model using four Monte Carlo Markov Chains each with 4000 warmup and 40000 sampling iterations. This was in order to obtain precise Bayes Factors using the Savage-Dickey ratio [@gronauBridgesamplingPackageEstimating2020].

>>>>

---
articletitle: |
  Self-talk interventions and sport/motor performance: An updated systematic review and Bayesian meta-analysis
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
author:
  - name: Hannah Corcoran
    affiliations:
      - ref: 1
  - name: James Steele
    affiliations:
      - ref: 1
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james.steele@solent.ac.uk
affiliations:
      - id: 1
        name: Department of Sport and Health, Solent University, UK
abstract: |
  Self-talk has been researched as an aspect of mental preparation for performance in sports and other motor tasks since the late 1980s. In 2011, Hatzigeorgiadis and colleagues systematically reviewed and meta-analysed three decades of self-talk intervention research including 32 studies. Yet, despite the general proliferation of meta-analyses, this topic has not been meta-analysed in the decade since their review which, at least for the main effect, provided a reasonably precise standardised mean difference estimate (0.48 [95% confidence interval: 0.38, 0.58]). Several further studies on self-talk interventions have been conducted in that time and it is of interest to explore what additional evidence they offer regarding the effects of self-talk interventions on sport/motor performance. Bayesian approaches are well positioned to explore how additional evidence changes our understanding of an effect; to see whether it has changed in sign, magnitude, or precision, or whether further research has largely been a ‘waste’. Therefore, our aim was to conduct an updated systematic review and Bayesian meta-analysis replicating the search, inclusion, and models of Hatzigeorgiadis et al. (2011). Informative priors were taken directly from Hatzigeorgiadis et al. A total of 34 studies providing 128 effects nested in 64 groups across experiments 42 were included in the final updated meta-analysis representing data from 18761 participants. The overall posterior pooled estimate for the standardised mean difference was almost exactly the same as the prior: 0.47 [95% quantile interval: 0.39, 0.56]. Bayes factors were calculated for a range of effect sizes and indicated that the included studies largely reflected ‘Weak’ evidence against effects ranging from 0.30 to 0.59, and only provided ‘Decisive’ evidence or greater against more extreme effects: either very small (i.e., <0.02) or large (i.e., >0.81). Results were largely similar for all moderator analyses which were updated too; either providing relatively weak evidence against effects found in the previous meta-analysis or evidence suggesting smaller effects for certain moderators. The findings of our updated Bayesian meta-analyses reiterate the positive effect of self-talk interventions on sport/motor performance. However, they also suggest that cumulatively the past decade and more of research has done little to further our understanding of these effects. Considering the limited resources and time for conducting research, it may be worth moving onto to other more pertinent questions regarding psychological constructs impact upon sport/motor performance.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [psychological intervention, cumulative evidence, research waste] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/dqwh5/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: references.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

@@

As we were interested in determining how much the new evidence produced since Hatzigeorgiadis et al. [-@hatzigeorgiadisSelfTalkSportsPerformance2011] had updated our understanding of the effects of self-talk interventions, we fit each model using four Monte Carlo Markov Chains each with 4000 warmup and 40000 sampling iterations. This was in order to obtain precise Bayes Factors using the Savage-Dickey ratio [@gronauBridgesamplingPackageEstimating2020]. Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions. These all showed good convergence with all $\hat{R}$ values close to 1 and  posterior predictive checks seemed appropriate distributions for the observed data (all diagnostic plots can be seen in the supplementary materials: [https://osf.io/xbcw8](https://osf.io/xbcw8)).
